---
title: "Comparison of Cylindar Count to Engine Displacement’"
format: pdf
editor: visual
---

## Introduction

Our client, Big Dawg, wants to know if having more cylinders in a car’s engine actually means a larger engine size. He thinks that more cylinders don’t always lead to a bigger engine. To find out, we’ll use data from the `mtcars` dataset in R to check the average engine size (displacement) for cars with different numbers of cylinders.

In this report, we’ll show the results in both a table and a chart. Using Quarto. This analysis will help Big Dawg see whether his belief about cylinders and engine size is accurate.

### 1. **Loading Necessary Libraries**

```{r}
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)

```

-   **ggplot2**: Used for creating the bar chart.

-   **dplyr**: Used for data manipulation, particularly grouping and summarizing data.

-   **knitr** and **kableExtra**: Used to create and style the table for displaying results.

### 2. **Loading the `mtcars` Dataset**

```{r}
data(mtcars)
head(mtcars)
```

```{r}
colnames(mtcars)
```

The column names of the `mtcars` dataset that we want are as follows:

1.  **mpg**: Miles per gallon (fuel efficiency)

2.  **cyl**: Number of cylinders

3.  **disp**: Engine size (displacement)

### 3. **Calculating the Mean Engine Size by Cylinder Count**

```{r}
mean_displacement <- mtcars %>%
  group_by(cyl) %>%
  summarise(
    mean_displacement = mean(disp, na.rm = TRUE)
  ) %>%
  arrange(cyl)

```

-   `group_by(cyl)`: Groups the data by the number of cylinders.

-   `summarise(mean_displacement = mean(disp, na.rm = TRUE))`: Calculates the mean engine size (`disp`) for each group (cylinder count). The `na.rm = TRUE` argument ignores any missing values.

-   `arrange(cyl)`: Orders the resulting data by cylinder count in ascending order.

The result, `mean_displacement`, is showing the average engine size for each cylinder count.

```{r}
mean_displacement %>%
  kable(
    col.names = c("No. of Cylinders", " Mean Size"),
    caption = "Mean Engine Size by Number of Cylinders"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    font_size = 12
  )

```

-   `kable()`: Creates a basic table with `mean_displacement` data. The `col.names` argument customizes column names, and `caption` adds a title.

-   `kable_styling()`: Adds style options like `striped` (alternating row colors), `hover` (highlight rows on hover), `condensed` (compact format), and `responsive` (adjusts size for different screen widths). `font_size` adjusts the table font size.

This results in a nicely formatted table showing mean engine size by cylinder count.
